---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

# Figures specification
* titled
* cited in numerical order in the text using Arabic numbers. 
* no additional charges for color. 
* submitted as individual files
* labeled with the Corresponding Author name, the appropriate figure number, and  orientation (e.g., “top”)
* Figures may be submitted as multipart panels.
*  *.tif* or .eps 
* final publication size; 1-column width or, if necessary, 1½ column width. The * 2-column width should not be used unless necessary.
* 1.0 column = 3.2 Inches, 8.2 Centimeters, 19.25 Picas
* 1.5 column = 4.5 Inches, 11.4 Centimeters, 27.00 Picas
* 2.0 column = 5.8 Inches, 14.8 Centimeters, 35.00 Picas
* height of all figures <= to 9.6inches / 24.5 cm / 58 picas.
* Calibri font sub sans serif Arial
* Multipanel labeled using uppercase 12-point Calibri Bold. 
* minimum text size of 6 points
* bold enough to be easily read after reduction, as should all symbols used in the * figure. Line or bar graphs or flow charts with text should be created in black and * white, not shades of gray, which are difficult to reproduce in even tones. If more * than two sets of data are represented, use of fill patterns or colors (not gray) is * suggested to present the data clearly.
* Minimum resolution is 300 dpi for color and grayscale figures, 600 dpi for * combination halftones, and 1000 dpi for line art.



```{r pkg}
library(tidyverse)
library(RColorBrewer)
library(readxl)
library(kableExtra)
library(grid)
library(patchwork)

```

```{r}

# palette
pal <- viridisLite::viridis(6)

# figure saving settings
units = "in"
fig_w <- 3.2
fig_w_1.5 <- 4.5
fig_h <- 4.5
fig_h_1.5 <- 4.5 * 1.5
dpi <- 300
device <- "tiff"

```

```{r import}
file <- here::here("data", "Meta-Analysis Data Collection.xlsx")

# added a row for column names more useful
# # skipped first two rows of of old col names

dat <- read_excel(file, sheet = "Data")

```


```{r tidy-data}
# recode country so those with one paper are
# "Other"
dat <- dat %>%
  mutate(Country2 = recode_factor(Country,
                                Brazil = "Brazil",
                                China = "China",
                                Denmark = "Denmark",
                                Egypt = "Egypt",
                                India = "India",
                                Japan = "Japan",
                                Korea = "Korea",
                                Netherlands = "Netherlands",
                                Poland = "Poland",
                                `Rep Korea` = "Rep Korea",
                                Russia = "Russia",
                                Spain = "Spain",
                                Taiwan = "Taiwan",
                                USA = "USA",
                                .default = "Other"))

dat <- dat %>%
  mutate(Phase_condensed = recode_factor(Phase,
                               I = "I",
                               `I/IIa` = "I",
                                II = "II",
                                IIa = "II",
                               IIb = "II",
                               III = "III"))

# test results - there is no "t" in the data file and there are more
# blanks values than t's in draft paper fig 6
dat <- dat %>%
  mutate(test_results = recode_factor(`Test Results?`,
                                       n = "Average value (n)",
                                       s = "Stated as 'standard phenotype' (s)",
                                       o = "Not discussed (o)",
                                       y = "Phenotypic markers reported (y)",
                                       t = "Tests done but not reported (t)"))
dat$test_results <- fct_relevel(dat$test_results,
                              "Phenotypic markers reported (y)", 
                              "Average value (n)",
                              "Tests done but not reported (t)",
                              "Stated as 'standard phenotype' (s)",
                              "Not discussed (o)")

# MOA - recode NS to Not stated
dat <- dat %>%
  mutate(MOA = recode_factor(MOA,
                             NS = "Not stated",
                             paracrine = "Paracrine",
                             immune = "Immune",
                             differentiation = "Differentiation",
                             multi = "Multiple"))

dat$MOA <- fct_relevel(dat$MOA,
                       "Paracrine",
                       "Immune",
                       "Differentiation",
                       "Multiple",
                       "Not stated")

# recode stem/stromal
dat <- dat %>%
  mutate(stem_stromal = recode_factor(`Stem/Stromal`,
                             `multipotent stromal` = "Multipotent Stromal",
                             other = "Other",
                             stem = "Stem",
                             stromal = "Stromal"))

dat$stem_stromal <- fct_relevel(dat$stem_stromal,
                                "Stromal",
                                "Stem",
                                "Multipotent Stromal",
                                "Other")

# Recode O, A and C into the 5 categories that appar 
dat <- dat %>%
  mutate(potency = ifelse(O == "y" &
                            A == "y" &
                            C == "y",
                          "Demonstrated tripotency",
                          ifelse(O == "n" &
                                   A == "n" &
                                   C == "n",
                                 "Not Stated",
                                 ifelse(O == "y" &
                                          A == "y" &
                                          C == "n",
                                        "Demonstrated bipotency (OA)",
                                        ifelse(O == "m"
                                               & A == "m" &
                                                 C == "n",
                                               "Claimed bipotency (OA)",
                                               ifelse(O == "m" &
                                                  A == "m" &
                                                  C == "m",
                                                  "Claimed tripotency", NA))))))        

dat$potency <- fct_relevel(dat$potency,
                                "Demonstrated tripotency",
                                "Demonstrated bipotency (OA)",
                                "Claimed tripotency",
                                "Claimed bipotency (OA)",
                                "Not Stated")                        


######### code for "Data Tab copy v1.csv"
# # column indices for markers
# firstmarker <- which(colnames(dat) == "CD73")
# lastmarker <- which(colnames(dat) == "SSEA.4")
# # vector of marker names
# markers <- names(dat[firstmarker:lastmarker])
# 
# datnum <- dat %>% 
#   select(markers) %>% 
#   mutate_all(as.numeric) 
# 
# # add the non marker columns
# notmarkers <- which(!(colnames(dat) %in% markers))
# datnum <- cbind(dat[notmarkers], datnum)
# 
# 
# dat$Stem.Stromal <- factor(dat$Stem.Stromal)
# dat$Year <- factor(dat$Year)
# datnum$Stem.Stromal <- factor(datnum$Stem.Stromal)
# datnum$Year <- factor(datnum$Year)


```

```{r}
# cd values for 

# note: file had missing value for Phas, this was looked up in "Meta-Analysis Data Collection.xlsx"
# as phase I/IIa
# note: ref 55, patient 6, marker CD45+/CD105+ contained value ND - I assumed this was no data.
# note - I have assumed range  is mean +/- se (not sd)
file <- here::here("data", "Paper 2 - Charts for Emma.xlsx")
cd_dat <- read_excel(file, sheet = "cd_values_single")
cd_dat2 <- read_excel(file, sheet = "cd_values_range")
```



```{r tidy-cd-data}
cd_dat <- cd_dat %>% 
  gather(key = marker, value = value, -Ref, -Phase, -patient, -notes)
cd_dat$value <- as.numeric(cd_dat$value)

cd_dat2 <- cd_dat2 %>% 
  gather(key = marker, value = value, -Ref, -Phase, -n)

for (i in 1:length(cd_dat2$value)) {
  temp <- strsplit(cd_dat2$value[i], split = "\\s.\\s") %>% unlist()
  cd_dat2$mean_value[i] <- as.numeric(temp[1])
  cd_dat2$se_value[i] <- as.numeric(temp[2])
}
cd_dat2 <- cd_dat2 %>% 
  filter(!is.na(value)) %>% 
  select(Ref, marker, mean_value, n, se_value, -Phase, -value)


```



# Figures

## Fig 1 - Article search funnel

```{r data-prep-fig1}
# create data
# short names for proof of method
orig <- data.frame(step_names = c("1.mesenchymal",
                                  "2.article",
                                  "3.cell tissue etc",
                                  "4.clinical etc",
                                  "5.not biotech etc",
                                  "6.exc conf etc"),
                   n = c(65706,
                         47031,
                         7814,
                         5502,
                         2055,
                         1986))

orig$step_names <- fct_relevel(orig$step_names,
                               "mesenchymal",
                               "article",
                               "cell tissue etc",
                               "clinical etc",
                               "not biotech etc",
                               "exc conf etc")

# prep polygons that connect
to_poly_orig <- orig %>%
  arrange(n) %>%
  mutate(
    group = row_number(),
    x1 = 0 + ((1 - n)/2),
    x4 = n + ((1 - n)/2),
    x2 = lag(x1), 
    x3 = lag(x4)) %>%
  mutate(
    x2 = ifelse(is.na(x2), x1, x2),
    x3 = ifelse(is.na(x3), x4, x3)
  ) %>%
  gather(xpoint, x, -step_names,-group, -n) %>%
  arrange(group, xpoint) %>%
  mutate(y = ifelse(xpoint %in% c("x2", "x3"), group, group + 1))


# EXAMPLE
# Create the placement of the stages in the plot,
#  by just adding a 0.5 to each row number


labels <- orig %>%
  arrange(n) %>%
  mutate(y = row_number() + 0.5)

# modify the labels
labels$step_names2 <- c( "EXCLUDE conference proceedings,\nretracted papers",
                         "EXCLUDE CATEGORIES (NOT\nbiotechnology applied microbiology oncology\nhaematology, engineering biomedical)",
                         "REFINE BY SOURCE TITLES\n(EXCLUDE NOT clinical/\ntranslation/regenerative medicine)",
                         "REFINE BY CATEGORY\ncell tissue engineering\nand transplantation",
                         "REFINE BY DOCUMENT TYPE\narticle",
                         "mesenchymal stem cells OR\nmesenchymal stromal cells OR\nMSC AND characterisation AND clinical")
```



```{r fig1}
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               size = 0) +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 2,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  labs(title = "Wilson, Alison, Figure 1. TOP" ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
    scale_fill_manual(values = rev(pal))

ggsave("figs/fig_1_Article_search_funnel.tif", 
       plot = article_funnel, 
       device = device,
       width = fig_w, 
       height = 3.0,
       units = units,
       dpi = dpi)
```

## Figure 3 Articles by country

```{r country-prep}
summary_country <- dat %>%
  group_by(Country) %>% 
  summarise(n = length(Country))


```

```{r fig3-country-bar}
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(Country, n), y = n)) +
  geom_bar(stat = "identity", fill = pal[3],
           size = 0) +
  xlab("") +
  scale_y_continuous(name = "Number of articles", 
                     breaks = seq(0, 16, 1),
                     minor_breaks = seq(0, 16, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 3. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 6, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()
ggsave("figs/fig_3_Article_country.tif", 
       plot = country_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
  
```
## Figure 4 Articles by Phase

### Six phase version

```{r phase-prep}
summary_phase <- dat %>%
  group_by(Phase) %>% 
  summarise(n = length(Phase),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
summary_phase <- summary_phase %>%
  mutate(Phase_condensed = recode_factor(Phase,
                               I = "I",
                               `I/IIa` = "I",
                                II = "II",
                                IIa = "II",
                               IIb = "II",
                               III = "III"))

summary_phase_condensed <- dat %>%
  group_by(Phase_condensed) %>% 
  summarise(n = length(Phase_condensed),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

```

```{r fig4-phase}
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(Phase,n), y = n)) +
  geom_bar(stat = "identity", fill = pal[3], size = 0) +
  xlab("") +
  geom_text(aes(y = n + 3, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 35, 5),
                     minor_breaks = seq(0, 35, 1),
                     limits = c(0, 35),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase.tif", 
       plot = phase_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```


### Three phase

```{r fig4-phase-3}
phase_3_bar <- summary_phase_condensed %>% 
  ggplot(aes(x = reorder(Phase_condensed, n), y = n)) +
  geom_bar(stat = "identity", fill = pal[3],size = 0) +
  xlab("") +
  geom_text(aes(y = n + 6, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 65, 5),
                     minor_breaks = seq(0, 65, 1),
                     limits = c(0, 65),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase_3.tif", 
       plot = phase_3_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```


## Figure 5 Articles by Source-Year

Done with bars side-by-side or stacked. 

### besides

```{r source_year_prep}
summary_tissue <- dat %>% 
  group_by(Source, Year = factor(Year), 
           .drop = FALSE) %>% 
  summarise(n = length(Source))

  
```

```{r fig5-source-year}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", size = 0, 
           position = position_dodge()) +
  scale_fill_manual(values = pal, 
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 12, 1),
                    minor_breaks = seq(0, 12, 1),
                    limits = c(0, 12),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year.tif", 
       plot = source_year, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

### stacked
```{r fig5-source-year-stack}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal,
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles",
                    breaks = seq(0, 18, 1),
                    minor_breaks = seq(0, 18, 1),
                    limits = c(0, 18),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year_stack.tif", 
       plot = source_year, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```


## Figure 6 Articles by Test quality

```{r test-quality-prep}
summary_test <- dat %>% 
  group_by(test_results) %>% 
  summarise(n = length(test_results))

summary_test$x <- 1
tot <- sum(summary_test$n)
n <- summary_test$n
summary_test$lab_y <- c(n[1]/2,
                        n[1] + n[2]/2,
                        n[1] + n[2] + n[3]/2,
                        n[1] + n[2] + n[3] + n[4]/2,
                        n[1] + n[2] + n[3] + n[4] + n[5]/2)
```

## Figure 7 Articles by Phase and Test quality reported

### Stacked by number of articles
```{r}

summary_phase_test <- dat %>%
  group_by(Phase_condensed, test_results) %>% 
  summarise(n = length(Phase))
```

```{r fig7-phase-test-n}
phase_test <- summary_phase_test %>% 
  ggplot(aes(x = Phase_condensed, y = n, fill = test_results)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal,
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 55, 10),
                     minor_breaks = seq(0, 55, 5),
                     limits = c(0, 55),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
#  labs(subtitle = "A") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 8),
        legend.position = "none")
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"))


```

### Stacked by percent of article

```{r fig7-phase-test-pc}
phase_test2 <- summary_phase_test %>% 
  ggplot(aes(x = Phase_condensed, y = n)) +
  geom_bar(aes(fill = test_results), 
           stat = "identity", 
           size = 0, 
           position = "fill") +
  scale_fill_manual(values = pal,
                    name = "Test results") +
  scale_y_continuous(name = "Proportion of Articles",
                     breaks = seq(0, 1, 0.2),
                #     limits = c(0, 1.2),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
  geom_text(data = summary_phase_condensed,
            aes(x = Phase_condensed, y = 1.05, vjust = 1,
                label = paste("n = ",n))) +
#  labs(subtitle = "B") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
      #  plot.title = element_text(hjust = 0, size = 8),
       legend.position = "none")



phase_test_both <- phase_test + phase_test2 + guide_area() +
  plot_annotation(title = "Wilson, Alison, Figure 7. TOP",
                  tag_levels = "A") +
  theme(legend.position = "top") +
  guides(fill = guide_legend(nrow = 2, title.position = "top"))


ggsave("figs/fig_7_characterisation_by_phase.tif", 
       plot = phase_test_both, 
       device = device,
       width = fig_w_1.5, 
       height = fig_h,
       units = units,
       dpi = dpi)
```

## Figure 9

```{r}
# data incoming
```

```{r}
# charc_marker %>% 
#   ggplot(aes(x = marker, y = n)) +
#   geom_bar(aes(fill = status), 
#            stat = "identity",
#            position = "fill",
#            colour = "white") +
#   scale_fill_manual(values = blues[4:7], name = "Status") +
#   scale_y_continuous(name = "Proportion of Articles") +
#   scale_x_discrete(expand = c(0, 0), name = "Marker") +
#   theme(panel.background = element_rect(fill = blues[2]),
#         panel.grid = element_blank(),)


```


## Relationship between MOA and stem/stromal

```{r}
moa_isct_summary <- dat %>% 
  group_by(MOA, stem_stromal) %>% 
  summarise(n = length(MOA))

res <- table(dat$stem_stromal, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)
test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))

```

```{r moa-isct}
moa_isct <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal, 
                    guide = guide_legend(nrow = 4, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x. TOP") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"))
ggsave("figs/fig_x_moa_isct.tif", 
       plot = moa_isct, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```
## Relationship between types of differentiation (OAC) and MOA

```{r}
table_info <- "There are five combinations for O, A and C potency. 'y' is demonstrated, 'm' is claimed but not demonstrated, 'n' is not mentioned. These were recoded into a single variable, 'potency'."


dat %>% 
  group_by(O, A, C, potency) %>% 
  summarise(n = length(O)) %>% 
  kable(caption = table_info) %>% 
  kable_styling() %>% 
  save_kable("figs/recode_differentiation_capacity.png")


```

```{r}
res <- table(dat$potency, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)

test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))
```


```{r moa-potency-prep}
moa_potency_summary <- dat %>% 
  group_by(potency, MOA) %>% 
  summarise(n = length(MOA))
```


```{r moa_potency}
moa_potency <- moa_potency_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = potency)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal, 
                    guide = guide_legend(nrow = 5, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x2. TOP") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"))
ggsave("figs/fig_x2_moa_potency.tif", 
       plot = moa_potency, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
```

```{r}
cd_dat_summary <- cd_dat %>% 
  group_by(Ref, marker) %>% 
  summarise(mean_value = mean(value),
            n = length(value),
            se_value = sd(value)/sqrt(n))
str(cd_dat_summary)
# length(unique(cd_dat$marker)) *length(unique(cd_dat$Ref))

# bind the range and the sinlge value dataframes
cd_dat_summary <- rbind(data.frame(cd_dat_summary), cd_dat2)

```


```{r}
# count the number of refs that have values
cd_refs_by_marker <- cd_dat_summary %>% 
  group_by(marker) %>%
  summarise(n_refs = length(mean_value[!is.na(mean_value)]))

cd_refs_by_marker$marker
isct <- c("Viability","CD73", "CD90", "CD105", "CD34", "CD45" , "CD11b", "CD14" , "CD79α", "CD19", "HLA-DR")
cd_refs_by_marker$ISCT <- "Other marker"
cd_refs_by_marker$ISCT[cd_refs_by_marker$marker %in% isct] <- "ISCT phenotype marker"

cd_refs_by_marker$ISCT <- fct_relevel(cd_refs_by_marker$ISCT,
                              "ISCT phenotype marker",
                              "Other marker")

```


```{r cd-refs-bar}
cd_refs_bar <- cd_refs_by_marker %>% 
  arrange(ISCT, n_refs) %>%
         mutate(marker = factor(marker, levels = marker) ) %>% 
  ggplot(aes(x = reorder(marker, ISCT), y = n_refs, fill = ISCT )) +
  geom_bar(stat = "identity",  size = 0) +
  xlab("") +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 15, 1),
                     minor_breaks = seq(0, 15, 1),
                     limits = c(0, 15),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_manual(values = pal[c(5, 2)], 
                    name = NULL) +
  labs(title = "Wilson, Alison, Figure X3. TOP" ) +
  annotate("segment", x = 0.5, xend = 10.5, 
           y = 14, yend = 14,
           colour = "black", size = 1) +
  annotate("segment", x = 0.58, xend = 0.58, 
           y = 14, yend = 13.5,
           colour = "black", size = 1) +
  annotate("segment", x = 10.38, xend = 10.38, 
           y = 14, yend = 13.5,
           colour = "black", size = 1) +
  annotate("text", x = 3,  y = 10, 
           label = "ISCT phenotype\nmarker", size = 4) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_x3_cd_refs_bar.tif", 
       plot = cd_refs_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}

cd_dat_summary_values <- cd_dat_summary %>% 
  filter(!is.na(mean_value))

cd_dat_summary_values$ISCT <- "Other marker"
cd_dat_summary_values$ISCT[cd_dat_summary_values$marker %in% isct] <- "ISCT phenotype marker"

cd_dat_summary_values$ISCT <- fct_relevel(cd_dat_summary_values$ISCT,
                              "ISCT phenotype marker",
                              "Other marker")
```

```{r}
cd_values_point_isct <- cd_dat_summary_values %>% 
  filter(ISCT == "ISCT phenotype marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x4_cd_values_isct.tif", 
       plot = cd_values_point_isct, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```
 
```{r}
cd_values_point_other <- cd_dat_summary_values %>% 
  filter(ISCT == "Other marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x5_cd_values_isct.tif", 
       plot = cd_values_point_other, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

