---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

# Figures specification
* titled
* cited in numerical order in the text using Arabic numbers. 
* no additional charges for color. 
* submitted as individual files
* labeled with the Corresponding Author name, the appropriate figure number, and  orientation (e.g., “top”)
* Figures may be submitted as multipart panels.
*  *.tif* or .eps 
* final publication size; 1-column width or, if necessary, 1½ column width. The * 2-column width should not be used unless necessary.
* 1.0 column = 3.2 Inches, 8.2 Centimeters, 19.25 Picas
* 1.5 column = 4.5 Inches, 11.4 Centimeters, 27.00 Picas
* 2.0 column = 5.8 Inches, 14.8 Centimeters, 35.00 Picas
* height of all figures <= to 9.6inches / 24.5 cm / 58 picas.
* Calibri font sub sans serif Arial
* Multipanel labeled using uppercase 12-point Calibri Bold. 
* minimum text size of 6 points
* bold enough to be easily read after reduction, as should all symbols used in the * figure. Line or bar graphs or flow charts with text should be created in black and * white, not shades of gray, which are difficult to reproduce in even tones. If more * than two sets of data are represented, use of fill patterns or colors (not gray) is * suggested to present the data clearly.
* Minimum resolution is 300 dpi for color and grayscale figures, 600 dpi for * combination halftones, and 1000 dpi for line art.



```{r pkg}
library(tidyverse)
library(RColorBrewer)
library(readxl)
library(kableExtra)
library(tiff)
library(grid)
library(patchwork)

```

```{r}

# palette
pal <- viridisLite::viridis(6)

# figure saving settings
units = "in"
fig_w <- 3.2
fig_w_1.5 <- 4.5
fig_w_2 <- 5.8
fig_h <- 4.5
fig_h_1.5 <- 4.5 * 1.5
dpi <- 300
device <- "tiff"

```

```{r import}
# file <- here::here("data", "Checked Data.xlsx")
# 
# # added a row for column names more useful
# # # skipped first two rows of of old col names
# 
# dat <- read_excel(file, sheet = "Checked Data")

```


```{r tidy-data}
# recode country so those with one paper are
# "Other"
# dat <- dat %>%
#   mutate(Country2 = recode_factor(Country,
#                                 Brazil = "Brazil",
#                                 China = "China",
#                                 Denmark = "Denmark",
#                                 Egypt = "Egypt",
#                                 India = "India",
#                                 Japan = "Japan",
#                                 Korea = "Korea",
#                                 Netherlands = "Netherlands",
#                                 Poland = "Poland",
#                                 `Rep Korea` = "Rep Korea",
#                                 Russia = "Russia",
#                                 Spain = "Spain",
#                                 Taiwan = "Taiwan",
#                                 USA = "USA",
#                                 .default = "Other"))
```

```{r}
# dat <- dat %>%
#   mutate(Phase_condensed = recode_factor(Phase,
#                                I = "I",
#                                `I/IIa` = "I",
#                                `I/II` = "I",
#                                 II = "II",
#                                 IIa = "II",
#                                IIb = "II",
#                                III = "III"))
```

```{r}
# dat <- dat %>%
#   mutate(test_results = recode_factor(`Test Results?`,
#                                        n = "Average value (n)",
#                                        s = "Stated as 'standard phenotype' (s)",
#                                        o = "Not discussed (o)",
#                                        y = "Phenotypic markers reported (y)",
#                                        t = "Tests done but not reported (t)"))
# dat$test_results <- fct_relevel(dat$test_results,
#                               "Phenotypic markers reported (y)", 
#                               "Average value (n)",
#                               "Tests done but not reported (t)",
#                               "Stated as 'standard phenotype' (s)",
#                               "Not discussed (o)")
```



```{r}
# MOA - recode NS to Not stated
# dat <- dat %>%
#   mutate(MOA = recode_factor(MOA,
#                              NS = "Not stated",
#                              paracrine = "Paracrine",
#                              immune = "Immune",
#                              differentiation = "Differentiation",
#                              multi = "Multiple"))
# 
# dat$MOA <- fct_relevel(dat$MOA,
#                        "Paracrine",
#                        "Immune",
#                        "Differentiation",
#                        "Multiple",
#                        "Not stated")
```

```{r}
# recode stem/stromal
# dat <- dat %>%
#   mutate(stem_stromal = recode_factor(`Stem/Stromal`,
#                              `multipotent stromal` = "Multipotent Stromal",
#                              regenerative = "Regenerative",
#                              stem = "Stem",
#                              stromal = "Stromal"))
# 
# dat$stem_stromal <- fct_relevel(dat$stem_stromal,
#                                 "Stromal",
#                                 "Stem",
#                                 "Multipotent Stromal",
#                                 "Regenerative")
# note: no longer an 'other' category, new regerative category
```

```{r}

# # Recode O, A and C into the 5 categories that appar 
# dat <- dat %>%
#   mutate(differentiation = ifelse(O == "y" &
#                             A == "y" &
#                             C == "y",
#                           "Demonstrated tripotency",
#                           ifelse(O == "n" &
#                                    A == "n" &
#                                    C == "n",
#                                  "Not Stated",
#                                  ifelse(O == "y" &
#                                           A == "y" &
#                                           C == "n",
#                                         "Demonstrated bipotency (OA)",
#                                         ifelse(O == "y" &
#                                                  A == "n" &
#                                                  C == "n",
#                                                "Demonstrated potency (O)",
#                                                ifelse(O == "m"
#                                                       & A == "m" &
#                                                         C == "n",
#                                                       "Claimed bipotency (OA)",
#                                                       ifelse(O == "m" &
#                                                                A == "m" &
#                                                                C == "m",
#                                                              "Claimed tripotency", NA)))))))        
# 
# dat$differentiation <- fct_relevel(dat$differentiation,
#                                 "Demonstrated tripotency",
#                                 "Demonstrated bipotency (OA)",
#                                 "Demonstrated potency (O)",
#                                 "Claimed tripotency",
#                                 "Claimed bipotency (OA)",
#                                 "Not Stated")                        
```



```{r}
# Recode O, A and C into 2 categories, all n = No, anything else = Yes 
# dat <- dat %>%
#   mutate(differentiation2 = ifelse(O == "n" &
#                                    A == "n" &
#                                    C == "n",
#                                  "No", "Yes"))      
# 
# dat$differentiation2<- fct_relevel(dat$differentiation2,
#                                 "No","Yes")                        
```
## write to file
```{r}
# file <- "data/checked_processed.csv"
# write_csv(dat, file)
```


# Read in the processed and corrected data

```{r}
file <- "data/checked_processed_aw_corrections.csv"
dat <- read_csv(file)
```

```{r}
dat <- dat %>%
  mutate(test_results = recode_factor(`Test Results?`,
                                       n = "Average value (n)",
                                       s = "Stated as 'standard phenotype' (s)",
                                       o = "Not discussed (o)",
                                       y = "Phenotypic markers reported (y)",
                                       t = "Tests done but not reported (t)"))
dat$test_results <- fct_relevel(dat$test_results,
                              "Phenotypic markers reported (y)",
                              "Average value (n)",
                              "Tests done but not reported (t)",
                              "Stated as 'standard phenotype' (s)",
                              "Not discussed (o)")
```

```{r}
# add a count of the number of markers tested for "Average value (n)". select columns like _test
test_cols <- names(select(dat, ends_with("_test")) )
dat <- dat %>%
  mutate(marker_count = rowSums(.[test_cols] == "p"))

```


```{r}

dat %>% group_by(MOA) %>% summarise(n = length(MOA))


dat$MOA <- fct_relevel(dat$MOA,
                       "Paracrine",
                       "Immune",
                       "Differentiation",
                       "Multiple",
                       "Not stated")
```


```{r}
dat %>% group_by(stem_stromal) %>% summarise(n = length(stem_stromal))

dat$stem_stromal <- fct_relevel(dat$stem_stromal,
                                "Stromal",
                                "Stem",
                                "Multipotent Stromal",
                                "Regenerative")
```


```{r}
# cd values for 

# note: file had missing value for Phas, this was looked up in "Meta-Analysis Data Collection.xlsx"
# as phase I/IIa
# note: ref 55, patient 6, marker CD45+/CD105+ contained value ND - I assumed this was no data.
# note - I have assumed range  is mean +/- se (not sd)
# file <- here::here("data", "Paper 2 - Charts for Emma.xlsx")
# cd_dat <- read_excel(file, sheet = "cd_values_single")
# cd_dat2 <- read_excel(file, sheet = "cd_values_range")
```



```{r tidy-cd-data}
# cd_dat <- cd_dat %>% 
#   gather(key = marker, value = value, -Ref, -Phase, -patient, -notes)
# cd_dat$value <- as.numeric(cd_dat$value)
# 
# cd_dat2 <- cd_dat2 %>% 
#   gather(key = marker, value = value, -Ref, -Phase, -n)
# 
# for (i in 1:length(cd_dat2$value)) {
#   temp <- strsplit(cd_dat2$value[i], split = "\\s.\\s") %>% unlist()
#   cd_dat2$mean_value[i] <- as.numeric(temp[1])
#   cd_dat2$se_value[i] <- as.numeric(temp[2])
# }
# cd_dat2 <- cd_dat2 %>% 
#   filter(!is.na(value)) %>% 
#   select(Ref, marker, mean_value, n, se_value, -Phase, -value)


```

```{r}
dat %>% group_by(O) %>% 
  summarise(n = length(O), pc = n/84)
```
```{r}
dat %>% group_by(A) %>% 
  summarise(n = length(A), pc = n/84)
```
```{r}
dat %>% group_by(C) %>% 
  summarise(n = length(C), pc = n/84)
```
```{r}
dat %>% group_by(differentiation) %>% 
  summarise(n = length(differentiation), pc = n/84)
```


# Figures

## Fig 1 - Article search funnel

```{r data-prep-fig1}
# create data
# short names for proof of method
orig <- data.frame(step_names = c("1.mesenchymal",
                                  "2.article",
                                  "3.cell tissue etc",
                                  "4.clinical etc",
                                  "5.not biotech etc",
                                  "6.exc conf etc"),
                   n = c(65706,
                         47031,
                         7814,
                         5502,
                         2055,
                         1986))

orig$step_names <- fct_relevel(orig$step_names,
                               "mesenchymal",
                               "article",
                               "cell tissue etc",
                               "clinical etc",
                               "not biotech etc",
                               "exc conf etc")

# prep polygons that connect
to_poly_orig <- orig %>%
  arrange(n) %>%
  mutate(
    group = row_number(),
    x1 = 0 + ((1 - n)/2),
    x4 = n + ((1 - n)/2),
    x2 = lag(x1), 
    x3 = lag(x4)) %>%
  mutate(
    x2 = ifelse(is.na(x2), x1, x2),
    x3 = ifelse(is.na(x3), x4, x3)
  ) %>%
  gather(xpoint, x, -step_names,-group, -n) %>%
  arrange(group, xpoint) %>%
  mutate(y = ifelse(xpoint %in% c("x2", "x3"), group, group + 1))


# EXAMPLE
# Create the placement of the stages in the plot,
#  by just adding a 0.5 to each row number


labels <- orig %>%
  arrange(n) %>%
  mutate(y = row_number() + 0.5)

# modify the labels
labels$step_names2 <- c( "EXCLUDE conference proceedings,\nretracted papers",
                         "EXCLUDE CATEGORIES (NOT\nbiotechnology applied microbiology oncology\nhaematology, engineering biomedical)",
                         "REFINE BY SOURCE TITLES\n(EXCLUDE NOT clinical/\ntranslation/regenerative medicine)",
                         "REFINE BY CATEGORY\ncell tissue engineering\nand transplantation",
                         "REFINE BY DOCUMENT TYPE\narticle",
                         "mesenchymal stem cells OR\nmesenchymal stromal cells OR\nMSC AND characterisation AND clinical")
```



```{r fig1}
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               size = 0) +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 2,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  labs(title = "Wilson, Alison, Figure 1. TOP" ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
    scale_fill_manual(values = rev(pal))

ggsave("figs/fig_1_Article_search_funnel.tif", 
       plot = article_funnel, 
       device = device,
       width = fig_w, 
       height = 3.0,
       units = units,
       dpi = dpi)
```

## Figure 3 Articles by country

```{r country-prep}
summary_country <- dat %>%
  group_by(Country) %>% 
  summarise(n = length(Country))


```

```{r fig3-country-bar}
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(Country, n), y = n)) +
  geom_bar(stat = "identity", fill = pal[3],
           size = 0) +
  xlab("") +
  scale_y_continuous(name = "Number of articles", 
                     breaks = seq(0, 16, 1),
                     minor_breaks = seq(0, 16, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 3. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 6, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()
ggsave("figs/fig_3_Article_country.tif", 
       plot = country_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
  
```
## Figure 4 Articles by Phase

### Eight phase version

```{r phase-prep}

summary_phase <- dat %>%
  group_by(Phase) %>% 
  summarise(n = length(Phase),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
summary_phase <- summary_phase %>%
  mutate(Phase_condensed = recode_factor(Phase,
                               I = "I",
                               `I/IIa` = "I",
                               `I/II` = "I",
                                II = "II",
                                IIa = "II",
                               IIb = "II",
                               III = "III",
                               NS = "NS"))

summary_phase_condensed <- dat %>%
  group_by(Phase_condensed) %>% 
  summarise(n = length(Phase_condensed),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

```

```{r fig4-phase}
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(Phase,n), y = n)) +
  geom_bar(stat = "identity", fill = pal[3], size = 0) +
  xlab("") +
  geom_text(aes(y = n + 4, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 1),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase.tif", 
       plot = phase_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```


### Three phase

```{r fig4-phase-3}
# phase_3_bar <- summary_phase_condensed %>% 
#   ggplot(aes(x = reorder(Phase_condensed, n), y = n)) +
#   geom_bar(stat = "identity", fill = pal[3],size = 0) +
#   xlab("") +
#   geom_text(aes(y = n + 6, label = pc) ) +
#   scale_y_continuous(name = "Number of Articles", 
#                      breaks = seq(0, 65, 5),
#                      minor_breaks = seq(0, 65, 1),
#                      limits = c(0, 65),
#                      expand = c(0, 0)) +
#   scale_x_discrete(expand = c(0, 0)) +
#   labs(title = "Wilson, Alison, Figure 4. TOP" ) +
#   theme_classic() +
#   theme(axis.title.y = element_blank(),
#         axis.text.x = element_text(size = 8, colour = "black"),
#         axis.ticks.y = element_blank(),
#         axis.text.y  = element_text(size = 8, colour = "black"),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         plot.title = element_text(hjust = 0, size = 4)) +
#   coord_flip()
# 
# ggsave("figs/fig_4_Article_phase_3.tif", 
#        plot = phase_3_bar, 
#        device = device,
#        width = fig_w, 
#        height = fig_w,
#        units = units,
#        dpi = dpi)

```


## Figure 5 Articles by Source-Year


```{r source_year_prep}
summary_tissue <- dat %>% 
  group_by(Source, Year = factor(Year), 
           .drop = FALSE) %>% 
  summarise(n = length(Source))
dat %>% 
  group_by(Source, 
           .drop = FALSE) %>% 
  summarise(n =length(Source), pc = n/84)

```



### stacked
```{r fig5-source-year-stack}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal[c(1,3,5,6)],
                    guide = guide_legend(nrow = 3, title = NULL )) +
  scale_y_continuous(name = "Number of Articles",
                    breaks = seq(0, 20, 2),
                    minor_breaks = seq(0, 20, 2),
                    limits = c(0, 20),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = c(0.2, 0.9),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"))
       
ggsave("figs/fig_5_Article_source_year_stack.tif", 
       plot = source_year, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```


## Figure 6 - three panels


### panel 1: Articles by Test quality

```{r test-quality-prep}
summary_test <- dat %>% 
  group_by(test_results) %>% 
  summarise(n = length(test_results),
            pc = paste0(round(100 * n / 84, 1 ), "%"))


```

```{r testquality}
test_bar <- summary_test %>% 
  ggplot(aes(x = reorder(test_results, n), y = n, fill = test_results)) +
  geom_bar(stat = "identity", size = 0) +
  xlab("") +
  geom_text(aes(y = n + 6, label = pc) ) +
  scale_fill_manual(values = pal,
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 50, 5),
                     minor_breaks = seq(0, 50, 1),
                     limits = c(0, 50),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0),
                   labels = c("Stated as 'standard\nphenotype' (s)",
                              "Phenotypic markers\nreported (y)",
                              "Tests done but\nnot reported (t)",
                              "Not discussed (o)",
                              "Average value (n)")) +
  labs(subtitle = "A") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4),
        plot.margin = margin(r = 0.1, b = 0.1, l = 0, unit="in")) +
  coord_flip()

# ggsave("figs/fig_6_Article_test_quality.tif", 
#        plot = test_bar, 
#        device = device,
#        width = fig_w, 
#        height = fig_w,
#        units = units,
#        dpi = dpi)

```

### panel 2 how many markers had average values  
```{r}
summary_marker_count <- dat %>% 
  filter(test_results == "Average value (n)") %>%
  group_by(marker_count) %>% 
  summarise(n = length(marker_count))


```

```{r markercount}
marker_bar <- summary_marker_count %>% 
  ggplot(aes(x = factor(marker_count), y = n)) +
  geom_bar(stat = "identity", fill = pal[2], size = 0) +
  scale_y_continuous(name = "Number of\n'Average value (n)' Articles ", 
                     breaks = seq(0, 12, 2),
                     minor_breaks = seq(0, 12, 2),
                     limits = c(0, 12),
                     expand = c(0, 0)) +
  scale_x_discrete(name = "Number of markers\nreported by an article",
                   expand = c(0, 0)) +
  labs(subtitle = "B") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.margin = margin(r = 0.2, t = 0,l = 0, unit="in")) 
```


### panel 3  Articles by Phase and Test quality reported (previously fig7 a)

### Stacked by number of articles
```{r}

dat %>%
  group_by(Phase_condensed) %>% 
  summarise(n = length(Phase_condensed),
            pc = n/84)

summary_phase_test <- dat %>%
  group_by(Phase_condensed, test_results) %>% 
  summarise(n = length(Phase_condensed))
```

```{r phasetest}
phase_test <- summary_phase_test %>% 
  ggplot(aes(x = Phase_condensed, y = n, fill = test_results)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal,
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 55, 10),
                     minor_breaks = seq(0, 55, 5),
                     limits = c(0, 55),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
 labs(subtitle = "C") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 8),
        legend.position = "none", 
        plot.margin = margin(r = 0.3, l = 0.1, t = 0, unit="in")) 

```

### Stacked by percent of article

```{r fig7-phase-test-pc}
# phase_test2 <- summary_phase_test %>% 
#   ggplot(aes(x = Phase_condensed, y = n)) +
#   geom_bar(aes(fill = test_results), 
#            stat = "identity", 
#            size = 0, 
#            position = "fill") +
#   scale_fill_manual(values = pal,
#                     name = "Test results") +
#   scale_y_continuous(name = "Proportion of Articles",
#                      breaks = seq(0, 1, 0.2),
#                 #     limits = c(0, 1.2),
#                      expand = c(0, 0)) +
#   scale_x_discrete(expand = c(0, 0), name = "Phase") +
#   geom_text(data = summary_phase_condensed,
#             aes(x = Phase_condensed, y = 1.05, vjust = 1,
#                 label = paste("n = ",n)),size=3) +
#  labs(subtitle = "B") +
#   theme_classic() +
#   theme(axis.text.x = element_text(size = 8, colour = "black"),
#         axis.text.y  = element_text(size = 8, colour = "black"),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         #  plot.title = element_text(hjust = 0, size = 8),
#         legend.position = "none",
#         plot.margin = margin(r = 0, l = 0, unit="in"))


```

```{r fig6}

layout <- "
aaaaaaa#
aaaaaaa#
aaaaaaa#
#bbbbccc
#bbbbccc
#bbbbccc
#bbbbccc
#bbbbccc
"


fig6 <- test_bar + marker_bar + phase_test + 
  plot_layout(design = layout) +
  plot_annotation(title = "Wilson, Alison, Figure 6. TOP")&
    theme(text = element_text(size = 8, hjust = 0))


ggsave("figs/fig_6.tif", 
       plot = fig6, 
       device = device,
       width = fig_w_2, 
       height = fig_h,
       units = units,
       dpi = dpi)

```


```{r}
dat
```




## Figure 7 Characterisation by marker.
```{r}

# select the test columns and count the number of articles for each marker and category
test_cols

charact_by_marker_ref <- dat %>% 
  select(ends_with("_test"), `Ref #`) %>% 
  pivot_longer(cols = -`Ref #`, names_to = "marker", values_to = "status")

# process the names to remove _test

charact_by_marker_ref <- charact_by_marker_ref %>% 
  mutate(marker2 = str_replace(marker, "_test", ""))

# change levels of status and markers.
# markers
# process test_col names
test_cols <- str_replace(test_cols, "_test", "")
# set the levels in that order
charact_by_marker_ref$marker2 <- factor(charact_by_marker_ref$marker2, levels = test_cols)

# status
# recode status
charact_by_marker_ref <- charact_by_marker_ref %>%
  mutate(test_status = recode_factor(status,
                             p = "Performed, value reported",
                             m = "Performed, no value reported",
                             n = "Not mentioned"))

charact_by_marker_ref$test_status <- fct_relevel(charact_by_marker_ref$test_status,
                                                     "Not mentioned",
                                                     "Performed, no value reported",
                                                     "Performed, value reported")




charact_by_marker_summary <- charact_by_marker_ref %>% 
  group_by(marker2, test_status) %>% 
  summarise(n = length(marker2))


```

```{r}
fig7 <- charact_by_marker_summary %>% 
  ggplot() +
  geom_bar(aes(x = marker2, y = n, fill = test_status),
           stat = "identity", width = 0.6) +
  scale_fill_manual(values =  pal[c(1, 2, 5)],
                    name = "Test results",
                    guide = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5 )) +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 90, 10),
                     minor_breaks = seq(0, 90, 10),
                     limits = c(0, 90),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0.5), name = "Characterisation attribute") +
  geom_rect(xmin = 1.5, xmax = 11.5,
            ymin = -Inf, ymax = Inf,
            size = 1, linetype = 1, 
            colour = "black", alpha = 0) +
  labs(title = "Wilson, Alison, Figure 7. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 6, colour = "black",
                                   angle = -60, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.2, "in"))
 
ggsave("figs/fig_7_charact_by_marker.tif", 
       plot = fig7, 
       device = device,
       width = fig_w_2, 
       height = fig_h,
       units = units,
       dpi = dpi)

 
```


```{r}
#alt fig
fig7alt <- charact_by_marker_ref %>% 
  ggplot() +
  geom_tile(aes(x = factor(`Ref #`), y = marker2 , fill = test_status),
            colour = "white")+
    scale_fill_manual(values =  pal[c(1, 2, 5)],
                    name = "Test results",
                    guide = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5 )) +
  scale_x_discrete(name = "Reference",
                     expand = c(0.01, 0)) +
  scale_y_discrete(expand = c(0, 0), name = "Characterisation attribute") +
  geom_rect(ymin = 1.5, ymax = 11.5,
            xmin = -Inf, xmax = Inf,
            size = 1, linetype = 1,
            colour = pal[6], alpha = 0) +
  labs(title = "Wilson, Alison, Figure 7ALT. TOP" ) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5, colour = "black", 
                                   angle = 90, hjust = 0, vjust = 0.3),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.2, "in"))
 
ggsave("figs/fig_7_alt.tif", 
       plot = fig7alt, 
       device = device,
       width = fig_w_2, 
       height = fig_w,
       units = units,
       dpi = dpi)
 
```




# Figure 8




# Figure 9


## Relationship between MOA and stem/stromal

```{r}
moa_isct_summary <- dat %>% 
  group_by(MOA, stem_stromal) %>% 
  summarise(n = length(MOA))


dat %>% 
  group_by(stem_stromal) %>% 
  summarise(n = length(stem_stromal),
            pc = n/84)

res <- table(dat$stem_stromal, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)
test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))

```

```{r moa-isct}
moa_isct <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal, 
                    guide = guide_legend(nrow = 4, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure 9. TOP") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black",
                                   angle = -60, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"))
ggsave("figs/fig_9_moa_isct.tif", 
       plot = moa_isct, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```


## Figure 10 assocation between differentiation2 and MOA
```{r}
diffmoasum <- dat %>% group_by(MOA, differentiation2) %>% 
  summarise(n = length(differentiation2))


```

```{r}
moa_diff <- diffmoasum %>% 
  ggplot(aes(x = MOA, y = n, fill = differentiation2)) +
  geom_bar(stat = "identity", size = 0) +
   scale_fill_manual(values = pal[c(2,5)], 
                    guide = guide_legend(nrow = 4, title = "Differentiation capacity"),
                    labels = c("Not demonstrated", "Demonstrated")) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure 10. TOP") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black",
                                   angle = -60, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"),
        legend.title = element_text(size = 8))
ggsave("figs/fig_10_moa_diff.tif", 
       plot = moa_diff, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
  
```



```{r}
res <- table(dat$differentiation2, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)
chisq.test(table(dat$differentiation2, dat$MOA))$exp;table(dat$differentiation2, dat$MOA)
test_anno <- paste0("There is  significant association between MOA and demonstration of differentiation capacity(", res$method, ": p = ", round(res$p.value, 3),") with more studies demonstrating differentiation for Immune MOA, and fewer for Paracrine and Multiple MOA than expected.")
```


```{r moa-potency-prep}
moa_potency_summary <- dat %>% 
  group_by(potency, MOA) %>% 
  summarise(n = length(MOA))
```


```{r moa_potency}
moa_potency <- moa_potency_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = potency)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal, 
                    guide = guide_legend(nrow = 5, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x2. TOP") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"))
ggsave("figs/fig_x2_moa_potency.tif", 
       plot = moa_potency, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
```

```{r}
cd_dat_summary <- cd_dat %>% 
  group_by(Ref, marker) %>% 
  summarise(mean_value = mean(value),
            n = length(value),
            se_value = sd(value)/sqrt(n))
str(cd_dat_summary)
# length(unique(cd_dat$marker)) *length(unique(cd_dat$Ref))

# bind the range and the sinlge value dataframes
cd_dat_summary <- rbind(data.frame(cd_dat_summary), cd_dat2)

```


```{r}
# count the number of refs that have values
cd_refs_by_marker <- cd_dat_summary %>% 
  group_by(marker) %>%
  summarise(n_refs = length(mean_value[!is.na(mean_value)]))

cd_refs_by_marker$marker
isct <- c("Viability","CD73", "CD90", "CD105", "CD34", "CD45" , "CD11b", "CD14" , "CD79α", "CD19", "HLA-DR")
cd_refs_by_marker$ISCT <- "Other marker"
cd_refs_by_marker$ISCT[cd_refs_by_marker$marker %in% isct] <- "ISCT phenotype marker"

cd_refs_by_marker$ISCT <- fct_relevel(cd_refs_by_marker$ISCT,
                              "ISCT phenotype marker",
                              "Other marker")

```


```{r cd-refs-bar}
cd_refs_bar <- cd_refs_by_marker %>% 
  arrange(ISCT, n_refs) %>%
         mutate(marker = factor(marker, levels = marker) ) %>% 
  ggplot(aes(x = reorder(marker, ISCT), y = n_refs, fill = ISCT )) +
  geom_bar(stat = "identity",  size = 0) +
  xlab("") +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 15, 1),
                     minor_breaks = seq(0, 15, 1),
                     limits = c(0, 15),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_manual(values = pal[c(5, 2)], 
                    name = NULL) +
  labs(title = "Wilson, Alison, Figure X3. TOP" ) +
  annotate("segment", x = 0.5, xend = 10.5, 
           y = 14, yend = 14,
           colour = "black", size = 1) +
  annotate("segment", x = 0.58, xend = 0.58, 
           y = 14, yend = 13.5,
           colour = "black", size = 1) +
  annotate("segment", x = 10.38, xend = 10.38, 
           y = 14, yend = 13.5,
           colour = "black", size = 1) +
  annotate("text", x = 3,  y = 10, 
           label = "ISCT phenotype\nmarker", size = 4) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_x3_cd_refs_bar.tif", 
       plot = cd_refs_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}

cd_dat_summary_values <- cd_dat_summary %>% 
  filter(!is.na(mean_value))

cd_dat_summary_values$ISCT <- "Other marker"
cd_dat_summary_values$ISCT[cd_dat_summary_values$marker %in% isct] <- "ISCT phenotype marker"

cd_dat_summary_values$ISCT <- fct_relevel(cd_dat_summary_values$ISCT,
                              "ISCT phenotype marker",
                              "Other marker")
```

```{r}
cd_values_point_isct <- cd_dat_summary_values %>% 
  filter(ISCT == "ISCT phenotype marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x4_cd_values_isct.tif", 
       plot = cd_values_point_isct, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```
 
```{r}
cd_values_point_other <- cd_dat_summary_values %>% 
  filter(ISCT == "Other marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x5_cd_values_isct.tif", 
       plot = cd_values_point_other, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

